---
title: "The Effects of COVID-19 on Police Issued Tickets"
subtitle: ""
author: 
  - Richard Guo
thanks: "The code and data supporting this analysis are available at: https://github.com/Richard-Guo1/Police_Issued_Tickets."
date: today
date-format: long
abstract: "To be included once analysis is completed"
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```


# Introduction

The categories of tickets are Aggressive Driving, All Compulsory Automobile Insurance Act transgressions (All CAIA), Distracted Driving, Other Highway Traffic Act infractions (Other HTA), and speeding. 


# Data {#sec-mydatasection}


```{r}
#| label: Tickets
#| fig-cap: Tickets
#| echo: false
  
data |> 
  ggplot(aes(x = `Offence Year`, weight = `Ticket Count`, 
             fill = `Offence Category`)) + 
  geom_bar() + labs(y = "Total Tickets")

```

Talk more about it.





# Discussion

## First discussion point {#sec-first-point}

## Second discussion point

## Third discussion point

## Weaknesses and next steps


\newpage

\appendix

# Appendix {-}


# Additional data details



\newpage


# References


